---
title: "athens_teaching_code"
author: "Argyris Stringaris"
date: "2023-12-06"
output: pdf_document
---


# A small tour in objects? Character vectors
```{r, echo=FALSE, message=FALSE,warning = FALSE}
my_name <- "Argyris"  # I just assigned a character Argyris, which always has to be in quotations to my_name. Now my_name is an # object saved in the memory of the computer.
my_name # if you run this you will see Argyris as an output
class(my_name) # class is a functionn, a command that asks the computer what kind of object it sees.
```

# A small tour in objects? Number vectors
```{r, echo=FALSE, message=FALSE,warning = FALSE}
my_numbers <- c(36,12,6)
my_numbers
```

# A small tour in objects. Lists
```{r, echo=FALSE, message=FALSE,warning = FALSE}
numbers_and_characters <- list(my_numbers, my_name) # list in this case is a command that tells R to create a list of the objects             #that are in the brackets. Lists are super useful and they can store pretty much any object.
numbers_and_characters
```

# Creating Data
We are now going to create in a gradual fashion a dataframe, which is R's way to call databases. 
Let's think of the following problem. 
You work at Aiginiteion Hospital in Athens. 
Let's say you want to measure depression in your patients there. 
And you want to compare those values of depression with the depression scores that you get in the community. 
You could go and collect data to find out--and eventually you may have to do it.
But before doing it, you want to prepare properly (this is very important--we will come back to its importance)
How do we prepare? 
By simulating. 

For the purposes of our study we are going to simulate data from 100 patients and another from 100 community-dwelling individuals

Here I tell the computer my number of subject **per group** 
```{r, echo=FALSE, message=FALSE,warning = FALSE}
n <- 100 
```


- exercise 1.1: vary this number 


Here I create my first data frame, that for Aiginiteion
```{r, echo=FALSE, message=FALSE,warning = FALSE}
aiginiteion_dep <- data.frame(
  ids = paste0("aiginiteion_",  1:n),
  dep_score = round(rnorm(n, 14, 2), 0)
)
```


```{r}
library(tidyverse)
aiginiteion_dep %>% 
  ggplot(aes(x = dep_score))+
  geom_histogram(position = "identity")
```