---
title: "Aiginiteion Teaching Evidence-Based Psychiatry"
author: "Argyris Stringaris"
date: "2023-12-06"
output: pdf_document
---

# First Hands-On Lecture on Coding and Basic Applied Statistics 

## Intro
These are the notes from the first hands-on lecture. 
The purpose is to get exposed to stats the proper way, i.e. by writing code. 
Learning to write code will equip you to solve many problems in your research and clinical practice. It can also be a sources of great fun.
Here we cover the basics. I have and will again refer you to standard works on the topic for a deeper understanding.
There are two main things you need to know about R. 
There are **objects** and there are **functions**
Think of functions as being active verbs and objects being nouns:functions do something with objects. 
There is also one more thing you need to know: the **assignment operator** which is symbolised as <- 
I will illustrate these things below.


## A small tour in objects? Character vectors
```{r}
my_name <- "Argyris"   
my_name 
class(my_name) 
```

- I just assigned, using <-, the string of characters "Argyris", which always has to be in quotations to my_name. Think of my_name as a box, a repository that can contain things like a string of characteris. After you run this, my_name is an # object saved in the memory of the computer.

- if you simply type _my_name_ and enter you will see Argyris as an output

- _class_ is a function that asks the computer what kind of object it sees. It is an inbuilt function, R has already got this and you can use it. Notice that the function can only run if you put the brackets around _my_name_. 


## A small tour in objects? Number vectors
You can assign numbers, as I show below. 
```{r}
my_numbers <- c(36,12,6)
my_numbers
```

- _my_numbers_ is now a number vector. Notice the = _c_ before the parentheses. It is a **function** and it means, combine the following elements (in this case numbers)

## Exercise 1.1: 
Use _c_ to create a vector that contains the names of you and two of your friendds. 


## A small tour in objects. Lists
Lists are also objects. They a very useful way of storing other objects. 
```{r}

numbers_and_characters <- list(my_numbers, my_name)

numbers_and_characters
```

-  list in this case is a function that tells R to create an object (in this case called _numbers_and_characters_) and to store in it a set of objects, namely those that are in the brackets. Lists are super useful and they can store pretty much any object.


## A small tour in objects. Dataframes
Dataframes are very important for anyone working with data. That's how datasets are generally stored. We will learn below how to do this.

## Creating Data
We are now going to create in a gradual fashion a dataframe, which is R's way to call databases. 
Let's think of the following problem. 
You work at Aiginiteion Hospital in Athens. 
Let's say you want to measure depression in your patients there. 
And you want to compare those values of depression with the depression scores that you get in the community. 
You could go and collect data to find out--and eventually you may have to do it.
But before doing it, you want to prepare properly (this is very important--we will come back to its importance)
How do we prepare? 
By simulating. 

For the purposes of our study we are going to simulate data from 100 patients and another from 100 community-dwelling individuals

## Number of subjects. 
Here I tell the computer my number of subject **per group** 
```{r, echo=FALSE, message=FALSE,warning = FALSE}
n <- 100 
```
- Here I just assigned a number to _n_

## My first simulated dataset
Here I create my first data frame, that for Aiginiteion
```{r}
set.seed(123)

aiginiteion_dep <- data.frame(
  ids = paste0("aiginiteion_",  1:n),
  dep_score = round(rnorm(n, 14, 2), 0)
)
```

- Let's break this down. Ignore the set.seed command for the moment, i will explain this below.

- Let's start with _data.frame_. This is a function that tells R to build a dataframe with the contents of what is in the brackets.

- The first thing I create is the ids. Every patient needs to have an id. Here I do a trick. I use another function (yes, in R you can use functions within a function) which is called _paste_ . As the name implies, paste puts together the word "aiginiteion" with the numbers that come out of 1:n. 

## Exercise 1.2. 
a) use your name and that of your friends from exercise 1 and use the paste command to have them appear next to each other. b) type the following: your year of birth (e.g.1992) : 2023. What do you see? Now type length(your year of birth (e.g.1992) : 2023)--what do you see? What is length?

- the second variable I create is the depression score. And here is where I use simulation. _rnorm_ is a function which generates _n_ **random** numbers from a normal distribution with the mean and standard deviation (sd) that we ask for. By having the size n, the mean and the sd, you can specify a normal distribution. In this case, I have specified n to be 100 (above) and gave it the mean of 14 and the sd of 2. I chose these numbers because one of my favourite depression questionnaires (the MFQ) had a mean of around 14 in my subject (though a broader sd). 

## Exercise 1.3
a) type ?rnorm into the console. What do you see on the right? b) type test <- rnorm(100,0,1) and then type test. What do you see? c) type hist(rnorm(100,0,1)) what do you see?
 
- finally, what is _set_seed_? It is a function that initialises pseudorandmo numbers in R. It helps us generate reproducible results when we use functions that use random numbers, such as rnorm. The following exercise will clarify this. 

## Exercise 1.4
a) Type set.seed(123) and generate test_1 <- rnorm(10, 0, 1). type set.seed(123) and generate test_2 <- rnorm(10, 0, 1). Type test_1 and test_2. What do you see? type test_1 == test_2, what do you get?
b)  type set.seed(456) and generate test_3 <- rnorm(10, 0, 1). Type test_3. What do you see? Try test_1 == test_3. What do you get? 
Note: the double = denotes logical equality, such as when you ask: "Argyris" == "Argyris" should come out as TRUE, whereas "Argyris" == "Kostas" should not. But "Argyris" != "Kostas" should also be TRUE. 



## Graphing my first simulated dataset
You now want to see how your data look like. 
```{r, echo=F, message=F}
library(tidyverse)

```


```{r, message = F}
aiginiteion_dep %>% 
  ggplot(aes(x = dep_score))+
  geom_histogram(position = "identity")
```

- you don't yet need to understand this code, but I will mention a few things for those of you interested. 

- library() invokes a particular bit of code that is a package of functions for doing things. In this case it calls an important superpackage that also allows me to draw the nice graphs below. 
 
- aiginiteion_dep is the dataset, the %>% operator is called a "pipe" and connects it with the next line which tells it how to plot the histogram. We will get t plotting. For the moment, use it as I have put it here. 

## Exercise 1.5
You have created a histogram. What does this mean? What is the y-axis, what is the x-axis, what does the height of each bar mean in this case? 


## Creating the control dataframe
As I said in the beginning, I want to compare patients in Aiginiteion to those in the community and test for differences. For this purpose, I am using the same process I have used above, namely create a dataset. 
```{r}
set.seed(123)
community_dep <- data.frame(
  ids = paste0("community_",  1:n),
  dep_score = round(rnorm(n, 8, 2), 0)
)
```

- Notice that I have used smaller mean and sd here as I expect these to be lower in the community. But I have used the same n

## Plotting the contorl dataframe
```{r, message = F}
community_dep %>% 
  ggplot(aes(x = dep_score))+
  geom_histogram(position = "identity")
```



## merge the patient and community samples 
Now you have the two datasets, you can put them together and see what happens.
```{r}
combined_df <- rbind(aiginiteion_dep, community_dep)
head(combined_df)
tail(combined_df)
dim(combined_df)

```

- in the first step, I basically stack the two dataframes on top of each other by using _rbind_, a command that says, bind rows. 

- you can check what rbind did by using the _head_ and _tails_ command. This allows you to see the beginning and end of the dataframe. You will see that aiginiteion is on top and community cases at the bottom. 


## Exercise 1.6
use the _rbind_ function as above, but flip around its contents, having community_dep first. What do you see when you use _head_ and _tail_? What do you see when you use _dim_?

- the _dim_ function tells you the dimensions of your dataframe. By conventions, dimensions are always reported as rows x columns. 

## Exercise 1.7 
change the _n_ further up in your code to n <- 1000. How many rows would you expect, how many columns? Why? Check using the _dim_function. (and put it back to n = 100)


## Adding a new variable to the dataset.
To make some of the subsequent analyses easier, I need to create a new variable that tells me which subjects come from the community and which from Aiginiteion.


```{r}
combined_df <- combined_df %>% 
  mutate(location = case_when(str_detect(combined_df$ids, "aiginiteion")~ "aiginiteion", 
                   str_detect(combined_df$ids, "community")~ "community") )

head(combined_df)
tail(combined_df)
```

- you don't need to understand everything here. We will explain this in future lessons. _mutate_ in this case creates the new variable _location_. And then I just use some text wrangling to say: whenever you see the word "aigniteion", e.g. "aiginiteion 10", in the _id_ column, put "aiginiteion" in the _location_ variable but whenever you see community, e.g. "community_19" in the _id_ column, put "community" in the _location variable_. 

## Exercise 1.8
What has changed in the dataset? What command would you run to answer the question?


## Plot the combined data!
Finally, I want to compare graphically the distributions of the two variables. 
```{r}
combined_df %>% 
  ggplot(aes(dep_score, fill = location))+
  geom_density( alpha = 0.5, bins = 100)
  
```

- Again, no need to explain what is going on with the code here, you will learn this in due course. 

- The main thing is to look at the graph. 

## Exercise 1.9
Try to interpret the graph. What is the y-axis, what is the x-axis. What does density mean? What is the darker area in the middle?


## The question you shouldn't rush to answer
Many of you will want to know whether the difference is statistically significant. 
Below is the answer. Feel free to try to understand what is going on below, looking at the output. 
I will try to convince you next time that there are better ways (for the time being) to think about differences and significance. But for the moment enjoy this.
```{r}
t.test(dep_score~ location, data = combined_df)
```

## save dataframe

## demonstrate how to use dataframe and each variable from it

## show how you can use it to do a simple comparison of means and sds





# Second Hands-On Lecture on Coding and Basic Applied Statistics 

## Intro
These are the notes from the second hands-on lecture. 

We will first review the exercises from last time. Your colleague **Giannis Rokas** has prepared the notes for it. 

After that, we will first learn how to 
- import data
- explore them 
- subset data
- estimate simple statistics by group
- write our first functions

**please work using a project. 
I have described how to do this in the email.

**please run first each line of code on its own so that you can get a feel of what is happening.

## Creating your first project
I will show you in the lesson how to create a project. This will be super helpful for you to work on data. 

## Saving a dataframe as a csv
```{r}
write.csv(combined_df, "aiginiteion_data.csv", row.names=FALSE)
```
- Here you save the dataset you created above. 

- Saving the data as a csv means you are using a universal and open source type file.

## Exercise 2.1
a) Save the dataset using a different name. Look under files on your right hand side in the project under "Files". How many new csv files do you see
b) as you will see, when you use this command, R saves the file in the current project location. You may however want to save it in another location. For that you will need a path. Google how to use a path for write.csv in R and apply this.


## Exploring the dataset
```{r}
aiginiteion_data <- read_csv("aiginiteion_data.csv") 

View(aiginiteion_data)

dim(aiginiteion_data)

head(aiginiteion_data)

tail(aiginiteion_data)

str(aiginiteion_data)
```

- The _read.csv_ function allows you to read the file into R.


## How to access the dataset
```{r}
aiginiteion_data$ids
aiginiteion_data$dep_score

mean(aiginiteion_data$dep_score)
sd(aiginiteion_data$dep_score)
```
 
- The _$_ sign is a function to access a named variable in a dataset
- _mean_ is a function as is _sd_

## Exercise 2.2
Let there be two vectors, x_1 and x_2 defined as follows
```{r}
x_1 <- c(1:10, 11, 12)

x_2 <- c(1:10, NA, 12)
```

a) run x_1 and x_2. What do you see? What does _NA_ stand for?
b) now use the _mean_ function to obtain the mean of each vector. What happens with x_2. Why? Why is it useful that R does this?
c) use ?_mean_ to find a solution to the problem and estimate the mean.


# How to subset the dataset
```{r}
aiginiteion_data[1, ]

aiginiteion_data[ ,1]

aiginiteion_data[1 ,1]

aiginiteion_data[aiginiteion_data$location == "aiginiteion" , ]


aiginiteion_data[aiginiteion_data$location == "aiginiteion" , ]$dep_score
aiginiteion_data[aiginiteion_data$location == "aiginiteion" , ]$ids
aiginiteion_data[aiginiteion_data$location == "community" , ]$dep_score


mean(aiginiteion_data[aiginiteion_data$location == "aiginiteion" , ]$dep_score)
mean(aiginiteion_data[aiginiteion_data$location == "community" , ]$dep_score)


```

- subsetting a dataset is one of the most important things to learn at this stage

- the tools above are like carving instruments that allow you to prepare the dataframe in exactly the way you like--in what is most interesting to you. 

- remember the convention _rows, columns_

## Exercise 2.3 
a) your boss is interested in the subject Giannis Papadopoulos and wants to know his depression score. You look him up and realise he corresponds to _id_ aiginition_55. What is his depression score? Please pull out only his depression score.

b) your boss is interested in a group of patients corresponding to the following aginiteion ids: 33, 44, 55, 66. how do you get those? you can use two ways. The tedious one makes use of the | , called the  _or_ operator, e.g. if you want to tell R to choose var_x == "a" | var_x == "b" it will look forany instance where either (or both) are tru. The more succinct way for multiple "or" statements is to use  _%in%_ , also called the vector memebership operator.



## How to loop
```{r}
for(i in 1:5){
  print("I love coding")
}

```

-  As discussed, the loop is one of the oldest and most revered way of doing iterations. 

## Exercise 2.4

- iterate over 1000 times


## Use a loop to find out the mean per group. 
See my comments next to the code.
Obviously, there are simpler ways of obtaining means and sds by group (see examples below). But loops equipe you with incredible flexibility for what you can do. Read the comments; however, the exercises will help you understand things best here.
```{r}
locations <- unique(aiginiteion_data$location) # run this line of code on its own.                         # what do you see? 

container_means <- 0 # here you create an empty vector to act as a container

container_sds <- 0 # same

for(i in 1: length(locations)){ # you are just saying: please make _i_ take all the numbers from 1 to length of locations. Run 1: length(locations) on its own in the console. What do you see?
  
container_means[i] <-  mean(aiginiteion_data[aiginiteion_data$location == locations[i],]$dep_score)  # decompose this. It is nothing than what you have done #above already in subsetting. Except here you are using the variable you created, #locations, and subset it too. in the first loop _i_  takes up the value of 1. In #the second one it takes on the value of 2.

container_sds[i] <-    sd(aiginiteion_data[aiginiteion_data$location == locations[i],]$dep_score)
 
}

container_means 
container_sds
```

## Exercise 2.5. 
It is always good to decompose code in its parts. 
a) run this piece of code on its own. What happens? What is the output, a vector or a dataframe
```{r}
aiginiteion_data[aiginiteion_data$location == locations[1],]

```

b) now run this piece of code. What comes out now?
```{r}
aiginiteion_data[aiginiteion_data$location == locations[1],]$dep_score
```


c) now run this piece of code. What comes out now?
```{r}
mean(aiginiteion_data[aiginiteion_data$location == locations[1],]$dep_score)
```

d) I have only replaced i with 1 above. Now run the above using 2. What do you see 


## other ways to obtain means by group. 
This is FYI at the moment. You can skip it for now.
```{r}

tapply(aiginiteion_data$dep_score, aiginiteion_data$location, mean)

aggregate(aiginiteion_data$dep_score, by = list(aiginiteion_data$location), mean)

library(tidyverse)
aiginiteion_data %>% 
  group_by(location) %>% 
  summarise(average_depression = mean(dep_score), sd_depression = sd(dep_score))


```

## Writing a function. 
Humanity has advanced through abstractions. 
Writing 10 = 5 + 5  refers to one possible case of y = x_1 + x_2 
The latter can also be written as f(x1, x2) = x_1 + x_2. It is a function with two variables, called a multivariable function. 
In R, x1 and x2 are called **arguments** of a function. 
They are abstractions, they can take on the values we subsequently tell them to. 
As in the case of f(x1, x2) = x_1 + x_2, where x_1 and x_2 can be any real number.
Follow the comments next to the code below.
```{r}

my_first_function <- function(date_birth, person_name){ # function simply tells R #to create a function with the arguments being _date_birth_ and _person_name_
age <- (Sys.Date() - as.Date(date_birth) ) # check what Sys.Date does. The function as.Date tells R to convert whatever date_birth contains into something that R can read as a date.

my_age <- paste(person_name, "is", round(age/365,3), "years old") # round is an #inbuilt function to round, I put _3_ as the number of decimals to round to. 


return(my_age) # this is very important. This is what you are asking the output of the function to be. 
}
my_first_function(c("1974-08-27", "2012-07-21"), c("Argyris", "Maria")) # I gave it two DOBs here and two names. run it. What comes out.


my_first_function(c("1993-4-27", "1993-07-26"), c("Giannis", "Lampros")) # this is what you gave me.


```



## My first step towards simulating power for statistics
The whole point of learning a language like R is so that you can test things that generic statistical software does not do, or does but in a non-transparent way. 
You also learn it so that you don't have to do things over and over again and waste time (and make easy-to-avoid mistakes).
Here we will create a function to do in only a line of code what we did above in several dozeens of lines: create a dataset. 
I am not doing it in the most economical way, but rather in the way that I think is didactically better. 
Remember the aiginiteion and the community datasets we created in our first lessons, out of which we created the combined one? If not go up and check--perhaps even copy and paste that code somewhere so you can see it as you work on the code below. 

```{r}

create_datasets <- function(name_gp1, name_gp2, n_gp1, n_gp2, mean_gp1, 
                            mean_gp2, sd_gp1, sd_gp2){

  #Here I just asked R to prepare a function that will take on all these #arguments. the ending gp refers to group, a generic name for any grouping
  #and I want it to have two groups 1 and 2, and each group to be able to accept
  # arguments for names (of the groups), n (i.e. sample size), and
  # a mean  and an sd (of the distribution of each sample)
  
  
df_1 <- data.frame(     # this is now very similar to the way I created the #aiginiteion or the community dataframe above, except I have made it be general 
  # and correspond to the arguments I have stipulated above. 
  
  ids = paste0(name_gp1, "_",  1:n_gp1),
  dep_score = round(rnorm(n_gp1, mean_gp1, sd_gp1), 0)
)


df_2 <- data.frame(   # same here.
  ids = paste0(name_gp2, "_",  1:n_gp2),
  dep_score = round(rnorm(n_gp2, mean_gp2, sd_gp2), 0))
  
  combined_df <- rbind(df_1, df_2) # now I only combine them.

combined_df <- combined_df %>%   # as in Lesson 1, no need to worry about this for # the moment, it is a trick I am using to get a third column in there easily.
  mutate(location = case_when(str_detect(combined_df$ids, name_gp1)~ name_gp1, 
                   str_detect(combined_df$ids, name_gp2)~ name_gp2) )

 

return(combined_df) # I want it to return me the datafram.

}

# OK we are done! Run now your first serious function. 
df_result <-  create_datasets("aiginiteion", "community", 250, 250, 14, 8, 4, 4)  
```

## Exercise 2.6 
Please run this function using the following argument paramters:
a) the same parameters as above, and assing to df_results_sd_4
b) the same parameters as above and SDs of 1, and assing to df_results_sd_1
c) the same parameters as above and SDs of 8, and assing to df_results_sd_8

## Plot the output of the function.
This allows you to depict the content of your dataframes in an interesting way
```{r}
plot_1 <- df_result %>% 
  ggplot(aes(dep_score, fill = location))+
  geom_density( alpha = 0.5, bins = 100)+
  xlim(-12, 30)
```

## Exercise 2.6 
Please run this function using the following argument paramters:
a) create a plot using the df_results_sd_4 data and assign it to the new object
plot_df_results_sd_4. Do the same with sd_1 and sd_8. 

b) you should now have three datasets. Run the following code after removing the #
```{r}
# plots_list <- list(plot_df_results_sd_1, plot_df_results_sd_4, plot_df_results_sd_8)
# 
# 
# ggsave("plots_list.pdf", plots_list,  device = "pdf")
```

c) Inspect the new pdf that comes out as a file under your Files bit. What do you notice? 


## Conducting my first simulation for power. Step 1. 
We had to go through this very quickly last time. 
We will revisit it step-by-step next time. 
If you want to have a look at it, it is the same as the code above. 
I have only made a little tweak. Instead of asking it to return the dataframes
that it creates, I have asked for the results of a t-test and specifically I pulled out the p-value from that test. 
I used a trick to do this, which you don't need to memorise and which is not important for now. But have a look.
```{r}
create_datasets <- function(name_gp1, name_gp2, n_gp1, n_gp2, mean_gp1, 
                            mean_gp2, sd_gp1, sd_gp2){

  
df_1 <- data.frame(
  ids = paste0(name_gp1, "_",  1:n_gp1),
  dep_score = round(rnorm(n_gp1, mean_gp1, sd_gp1), 0)
)


df_2 <- data.frame(
  ids = paste0(name_gp2, "_",  1:n_gp2),
  dep_score = round(rnorm(n_gp2, mean_gp2, sd_gp2), 0))
  
  combined_df <- rbind(df_1, df_2)

combined_df <- combined_df %>% 
  mutate(location = case_when(str_detect(combined_df$ids, name_gp1)~ name_gp1, 
                   str_detect(combined_df$ids, name_gp2)~ name_gp2) )

 my_test <- t.test(dep_score~ location, data = combined_df)

return(my_test$p.value)

}

```


## Here comes the actual iteration. This is for next time. 
```{r}
df_result <- replicate(100, create_datasets("aiginiteion", "community", 250, 250, 14, 12, 7, 7))
```



# Third Hands-On Lecture on Coding and Basic Applied Statistics 

## Summary  
We mainly went through the exercises from last time that Nikos had prepared. 
I have the following exercises for next time. 


## Exercise 3.1  

write a function that takes a dataframe as an argument and returns a dataframe containing the n (number of observations), mean, standard deviation, median and inter-quartile ranges for specified columns of the dataframe.
Please write your own function; you may use  inbuilt ones as you do so (e.g. mean()), but it would be idea if you created them yourself.

## Exercise 3.2 

Modify the function you created so that it can be applied across different groups, e.g. the mean of group a and group b.

## Exercise 3.3 

Please import the dataset for_athens_teaching_hands_on_exercises_3.csv 
Describe its dimensions and structure of its variables. 
This dataframe contains all the data from RCTs in either psychotherapy or medication trials and has both arms, active and controls.

## Exercise 3.4  

Use your function on the  cohens_d variable of the dataframe for_athens_teaching_hands_on_exercises_3, across the four_level_variable.

## Exercise 3.5 (optional)  

Graph this realationship, ideally encoding the size of the study (baseline_n) in the graph (e.g. as bubble size). 

## Exercise 3.6  

Write a short report of your findings. 



####################### END HERE #######################

## write a function to estimate means
```{r}
calculate_group_means <- function(df, var_for_group, var_outcome) {
  
  # Convert the grouping variable to character type
  df[[var_for_group]] <- as.character(df[[var_for_group]])
  
  # Get unique levels of the grouping variable
  group_levels <- unique(df[[var_for_group]])
  
  # Initialize an empty vector to store means
  means <- 0
  
  # Loop over each group
  for (i in seq_along(group_levels)) {
    group <- group_levels[i]
    
    # Subset the dataframe for the current group
    subset_data <- df[df[[var_for_group]] %in% group, ]
    
    # Calculate the mean of the outcome variable for the current group
    means[i] <- mean(subset_data[[var_outcome]], na.rm = TRUE)
  }
  
  # Create a data frame with group levels and corresponding means
  result <- data.frame(
    Group = group_levels,
    Mean = means
  )
  
  return(result)
}

# Example usage:
# Assuming 'df' is your dataframe, 'group_var' is your grouping variable,
# and 'outcome_var' is your continuous outcome variable


result <- calculate_group_means(aiginiteion_data, "location", "dep_score")

print(result)
```



```{r}
f <- function(x) c(mean(x), sd(x))
tapply(aiginiteion_data$dep_score, aiginiteion_data$location, f)
```




## 4. General points about regression

We discussed last time some of the fundamental maths, e.g. partial derivatives for the least squares. 
Today, we are going to start with a very simple problem about predicting. 


```{r}
library(tidyverse)
df_height_weight_all <- read.csv("~/teaching_athens/df_height_weight.csv")


df_height_weight_all$height_cm <- df_height_weight_all$Height.Inches*2.54

df_height_weight_all$weight_kg <- df_height_weight_all$Weight.Pounds.*0.45359237


df_height_weight_all  <- df_height_weight_all %>% 
  dplyr:: select(-c(Height.Inches., Weight.Pounds.))


#what did I do here?
df_height_weight_train <- df_height_weight_all[sample(nrow(df_height_weight_all), nrow(df_height_weight_all)*0.1),]





df_height_weight_train %>% 
  ggplot(aes(height_cm, weight_kg))+
  geom_point()


overall_model <- lm(weight_kg  ~ height_cm, df_height_weight_train  )
summary(overall_model)

overall_model_predictor_scaled <- lm(weight_kg  ~ scale(height_cm), df_height_weight_train  )
summary(overall_model_predictor_scaled)

overall_model_fully_scaled <- lm(scale(weight_kg)  ~ scale(height_cm), df_height_weight_train  )
summary(overall_model_fully_scaled)

cor.test(df_height_weight_train $weight_kg, df_height_weight_train $height_cm)

predicted_values <- predict (overall_model)

df_pred_actual <- data.frame(predicted_values = predicted_values, 
                             actual_values = df_height_weight_train $weight_kg)


head(df_pred_actual)


# what is the mean absolute error
df_pred_actual$abs_error <- abs(df_pred_actual$predicted_values - df_pred_actual$actual_values)
mean(df_pred_actual$abs_error)


#find the mean squared error


# find the R_squared

# the plot in base R
plot(x = predict (overall_model), y = df_height_weight_train$weight_kg)
abline(a = 0,                                        
       b = 1,
       col = "red",
       lwd = 2)


# create test data
df_test <- setdiff(df_height_weight_all, df_height_weight_train)

dim(df_test)

df_test <- df_test[sample(nrow(df_test), nrow(df_test)*0.01),]

dim(df_test)

out_of_sample_pred_values <- predict(overall_model, df_test)


#MAE 
mean(abs(df_test$weight_kg - out_of_sample_pred_values))


#### Do k-fold cross-validation


num_folds <- 5

num_rows <- nrow(df_height_weight_train)


fold_indices <- sample(rep(1:num_folds, length.out = num_rows))


cv_abs_errors <- 0

# Perform 5-fold cross-validation
for (i in 1:num_folds) {
  
  # each test dataset
  fold_data <- df_height_weight_train[fold_indices == i, ]
  
  # the training set
  train_data <- df_height_weight_train[fold_indices != i, ]
  

  lm_model <- lm(weight_kg  ~ height_cm, data = train_data)
  

  predictions <- predict(lm_model, newdata = fold_data)
  

  cv_abs_errors[i] <- mean(abs(fold_data$weight_kg - predictions))
}

# Calculate the average cross-validation error
average_cv_abs_error <- mean(cv_abs_errors)

average_cv_abs_error


```

